---
title: "Michael Hornsby EDRM 878 Assignment 1 Notebook"
output: html_notebook
---

***

##### Grading comments

**Nice work! You may want to take a look at my work. After reading your**
**conclusions, it wasn't obvious to me that you recognized why the time was**
**quicker for building cylinder-based towers. Looking at the analysis for**
**both time and for numbers, it was clear to me that the time was longer**
**with cubes because the children built higher towers. Also, some of the**
**older children took substantially longer because they built much higher**
**towers.**

**Score: 20/20**

***


```{r include=FALSE}
library(here)
library(tidyverse)
library(ggrepel)
library(scales)
library(readxl)
library(purrr)
library(dplyr)
library(car)
library(GLMsData)
```


Problem Set 1.9

Part 1)

```{r}
data(blocks)

blocks$Trial <- as.factor(blocks$Trial)
blocks$Shape <- as.factor(blocks$Shape)

summary(blocks)
```



Part 2)

```{r}
blocks %>% 
  ggplot(aes(x = Shape, y = Time)) +
  geom_boxplot() +
  labs(title = "Boxplots of Time by Block Shape",
       x = "Block Shape")
```

```{r}
blocks %>% 
  ggplot(aes(x = Trial, y = Time)) +
  geom_boxplot() +
  labs(title = "Boxplots of Time by Trial")
```

```{r}
blocks %>% 
  ggplot(aes(x = Age, y = Time)) +
  geom_point() +
  labs(title = "Scatterplot of Time by Age")
```



Part 3

Looking at the boxplots of time taken by block shape, it appears as though
using cube blocks resulted in an overall longer time taken. 
I think I would have expected the opposite.  I would have guessed children
would be quicker at stacking cubes than cylinders.
Looking at the boxplots of time by trial, it looks as though children took
about the same time for both trials.  As a general trend in the scatterplot,
there doesn't seem to be an effect of age on time.  If I ignore those 2 points
greater than a time of 125, there doesn't seem to be an significant
relationship that I can visualize form the scatterplot.



Part 4

```{r}
blocks %>% 
  ggplot(aes(x = Shape, y = Number)) +
  geom_boxplot() +
  labs(title = "Boxplots of Number of Blocks by Block Shape",
       x = "Block Shape",
       y = "Number of Blocks")
```

```{r}
blocks %>% 
  ggplot(aes(x = Trial, y = Number)) +
  geom_boxplot() +
  labs(title = "Boxplots of Number of Blocks by Trial",
       y = "Number of Blocks")
```

```{r}
blocks %>% 
  ggplot(aes(x = Age, y = Number)) +
  geom_point() +
  labs(title = "Scatterplot of Number of Blocks by Age",
       y = "Number of Blocks")
```



Part 5)

Looking at the boxplots of Number by Shape, it appears as though children using
cubes used more blocks than children using cylinders on a fairly consistant
basis.  Children at trial 2 appear to use slightly more blocks than at trial 1,
although, I wouldn't think this effect would be significant at any reasonable
alpha level looking at the boxplots.
Looking at the scatterplot of number of blocks by age, there does appear to be
a slightly positive linear relationship between number of blocks and age.  As
age increases, the number of blocks used on average appears to increase
slightly.



Problem Set 2.18)

Part 1)

Block shape does appear to have an effect on time taken.  As stated above, it 
looks as though children using cylinders took less time overall than children
using cubes, although it isn't apparent immediately to me if the effect will be
significant at any reasonable alpha level.  The two block shape groups do not
appear to really have equal variance nor do they look exactly normal given
their visible upward skew.  Trial does not appear to affect time at all when
looking at the boxplots.  It isn't apparent to be me immediately when looking
at the scatterplot of time by age whether age will have any significant effect.
However, even if a significant effect of age on time is found, I would wonder
whether or not that effect would still be present if those 2 times greater than
125 didn't have enough weight to create an effect.  Those two points aside, I
wouldn't be on any significant age effect.



Part 2)

If we assume that age, shape, and the interaction effect of age and shape are
significant and provide an appropriate model for the data, we can conclude a
few things.  The amount of time taken significantly changes with age.  The
amount of time taken also significantly differs between the two shapes.
Additionally, there is a significant interaction between age and shape such
that as children age they become 'proficient' in stacking blocks of one
type at a faster rate than they do of the other type. (I am assuming that
children will not become less proficient at stacking blocks, although
negative relationships and interactions are possible.)



Part 3)

If we assume that age, trial, and the interaction effect of age and trial are
significant and provide an appropriate model for the data, we can conclude a
few things.  The amount of time taken significantly changes with age.  The
amount of time taken also significantly differs between the two trials.
Additionally, there is a significant interaction between age and trial such
that the slope of time by age differs significantly between the two trials.



Part 4)

This systematic component, given all effects are significant, claims that
age, shape, trial, the interaction of age and trial, and the interaction of
shape and trial are significant. (Or, depending on the effects and 
significance, could just mean the interactions may be significance. I know
that is a whole debate by itself.)  If we assume all effects are significant, 
we know that time differs significantly with age, by shape, and by trial.
Additionally, the slope of time by age differs for block shape AND for trial.



Part 5)

```{r}
anova(lm(Time ~ Shape, data = blocks))
```

We see here that time does differ significantly between block shapes at 
any reasonable alpha level with a p-value of 0.004315.



Part 6)

```{r}
anova(lm(Time ~ Age * Shape, data = blocks))

anova(lm(Time ~ Age, data = blocks))
```

When looking at these two models, we can see that we cannot conclude that
older children take less time when accounting for block type.  Additionally,
we cannot conclude that older children take less time to build the towers at all.
The p-values of age and the p-value of the age:shape interaction are not
significant in either model at any reasonable alpha level.  As confirmed in
the scatterplot above, age is not a significant predictor of time taken.



Part 7)

```{r}
anova(lm(Time ~ Shape, data = blocks))
```

When predicting time taken to the build the towers, only the shape of the
blocks appears to be a significant predictor.  Age and trial are not
significant or produce any significant interactions even with shape.
Using shape as the only predictor of time is significant, and produces a
suitable final model.



Part 8)

The significant predictor of time taken to build the towers is the shape
of the blocks used.  The shape variable has a p-value of 0.004315, which is
significant at any reasonable alpha level.  We can conclude that children using
cylinders take significantly less time to build the towers than children using
cube blocks on average.