---
title: "Michael Hornsby EDRM 878 Final"
output: html_notebook
---

libraries:

```{r include=FALSE}
library(here)
library(tidyverse)
library(ggrepel)
library(scales)
library(readxl)
library(purrr)
library(dplyr)
library(car)
library(GLMsData)
library(MASS)
library(statmod)
```


NORMAL:

Description:

Mean-Variance Relationship:

  The variance of mu can be treated as 1 here since the variance is captured in the
    dispersion parameter sigma squared, and the variance of a normal distribution is
    simply sigma squared

Dispersion:

  The dispersion is sigma squared - the variance of a normal distribution,
    this will have to be estimated

Potential Link Functions:
Methods of Inference:
Saddlepoint Approximation and Central Limit Theorem:

  The saddlepoint approximation is exact for the normal distribution.
  The central limit theorem is exact for the normal distribution as well.

When to adjust for overdispersion:
What type of residual to use:
Diagnostics:
Scale transformations to use:
Primary effect size estimates:



Analysis:

```{r}
data(crawl)

summary(crawl)
```

```{r}
crawl %>% 
  ggplot(aes(x = Temp, y = Age)) +
  geom_point() +
  labs(title = "Scatterplot of Age on Temp",
       y = "Age at which baby started to crawl",
       x = "Monthly average temperature six months after birth month")

crawl %>% 
  ggplot(aes(y = Temp)) +
  geom_boxplot() +
  labs(title = "Boxplot of Temp")

crawl %>% 
  ggplot(aes(y = Age)) +
  geom_boxplot() +
  labs(title = "Boxplot of Age")

crawl %>%
  ggplot() +
  geom_qq(aes(sample = Temp)) +
  labs(title = "QQ-plot for Temp")

crawl %>%
  ggplot() +
  geom_qq(aes(sample = Age)) +
  labs(title = "QQ-plot for Age")

```

Well, here we can see from the scatterplot that things look mostly linear.  If I could
remove that one observation around Temp 50 with the lowest Age, I would be extremely happy just using the two variables as they are.  The boxplot for temp looks really good and mostly normal with a small skew.  The boxplot for age looks as though it has much larger skew.  The qq-plots mirror this.  I'm going to try a ladder of powers with Age to see what I get and see if any transformation makes things better.

```{r}
boxcox(Age ~ Temp,
       lambda = seq(-5.00, 20.00, by = .05), data = crawl)
```

Oh boy, this doesn't look good.  Both our predictor and response show slight skew, and based on
all of my exploration trying to transform the response, I think it is best to leave it as is.
Twelve data points isn't a lot and I wouldn't expect a miracle from the model.

```{r}
normal.model <- glm(Age ~ Temp, data = crawl)
summary(normal.model)
confint(normal.model)
```

Ok, all things considered, this model looks nice.  We have a significant p-value of 0.0113
for Temp, which is significant at an alpha = 0.05 level.  We have an Intercept age of
35.67806 and a Temp coefficient of -0.07774.  Thus our model is:

Age = 35.67806 - 0.07774(Temp)

This model will give us the expected Age at which the baby will start to crawl given the
monthly average temperature six months after the birth month of the baby.

We have our 95% confidence interval for the Intercept:
  (33.0957562, 38.26036294)
And our 95% confidence interval for the Temp coefficient:
  (-0.1269348, -0.02854351)


```{r}
resid.raw <- resid(normal.model)

fit.values <- fitted(normal.model)

plot(fit.values, resid.raw)

influence.measures(normal.model)
```

As far as diagnostics go, the residual plot looks to have an even enough spread with the exception of that one point.  R marks 3 obserations as highly influential.  I wouldn't be 
concerned with observations 2 and 3.  They have low cook's D.  However, again, that one
observation that seems to be an outlier has a significantly higher cook's D relative
to the rest of the observations.  I think I would suggest returning to the theory or
researcher on how to address this point.  The model holds when including it, but removing
it may be acceptable.  Also, more than 12 data points would be nice as well.  However,
regardless of this highly influential point, we can safely conclude that Temp does
significantly predict Age in the context of these variables.

Plot of model relationship:

```{r}
crawl %>% 
  ggplot(aes(x = Temp, y = Age)) +
  geom_point() +
  labs(title = "Scatterplot of Age on Temp",
       y = "Age at which baby started to crawl",
       x = "Monthly average temperature six months after birth month") +
  geom_smooth(method = 'lm', se = FALSE)
```



Additional:

Our Wald test is the same as above:

```{r}
summary(glm(Age ~ Temp, data = crawl))
```

The score test:

```{r}
crawl.m0 <- glm(Age ~ 1, data = crawl)
z.stat <- glm.scoretest(crawl.m0, crawl$Temp)
p.val <- 2 * pnorm(abs(z.stat), lower.tail = FALSE)
round(c(score.stat = z.stat, P = p.val), 4)
```


The likelihood ratio test:

```{r}
fit <- glm(Age ~ Temp, data = crawl)
as.data.frame(anova(fit, test = "Chisq"))
```

Well, all 3 tests suggest that Temp does have a significant effect on Age. The
p-values differ, especially for the likelihood ratio test, but I think this is most likely because of the smaller sample size.







BINOMIAL:

Description:

Mean-Variance Relationship:

  For mean mu, the variance will be:
    V(mu) = mu(1 - mu) where mu = np

Dispersion:

  The dispersion is 1/m where m in the number of trials

Potential Link Functions:
Methods of Inference:
Saddlepoint Approximation and Central Limit Theorem:

  The saddlepoint approximation for the binomial will be sufficient when
    min(my) >= 3 and min(m(1-y)) >= 3
  The central limit theorem will be sufficiently accurate for the binomial when
    min(my) >= 5 and min(m(1-y)) >= 5

When to adjust for overdispersion:
What type of residual to use:
Diagnostics:
Scale transformations to use:
Primary effect size estimates:




Analysis:

```{r}
data(belection)

belection$totalran = belection$Females + belection$Males
belection$propfemale <- belection$Females / (belection$total)
belection$Region <- as.factor(belection$Region)
belection$Party <- as.factor(belection$Party)

summary(belection)
```

We're going to be using a binomial EDM with a logit link and we have region and
party to explore.

```{r}
binom.logit <- glm(Females/totalran ~ Party,
                data = belection,
                family = binomial,
                weights = totalran)

summary(binom.logit)
anova(binom.logit, test = "Chisq")
confint(binom.logit)
```

Well, after factoring in the weights and our 2 factor predictors, Party and Region, I settled
on the model above.  The model is a weighted logistic regression using only Party
as a predictor.  Party has a significant p-value of 4.152 x 10^(-11), which is
significant at any reasonable alpha level.  In terms of the parties, this model results
in the following expected proportions of females running:

Conservative: 0.0899
Green: 0.2381
Labour: 0.1987
LibDem: 0.2152
Other: 0.1982

as shown below:

```{r}
exp(-2.3148)/(1 + exp(-2.3148))
exp(-2.3148 + 1.1516)/(1 + exp(-2.3148 + 1.1516))
exp(-2.3148 + 0.9206)/(1 + exp(-2.3148 + 0.9206))
exp(-2.3148 + 1.0209)/(1 + exp(-2.3148 + 1.0209))
exp(-2.3148 + 0.9174)/(1 + exp(-2.3148 + 0.9174))
```

The 95% confidence intervals for these coefficients are noted above.

Let's look at diagnostics:
```{r}
resid.raw <- resid(binom.logit)

fit.values <- fitted(binom.logit)

plot(fit.values, resid.raw)

influence.measures(binom.logit)
```

I looked through all of the points marked as influential, but they all look ok.  Cook's D's
all look good, and the residuals show a nice, mostly even spread about 0.  
I think this is a nice model altogether.  I'm actually big on following
politics including UK politics, and this seems to follow the trend I would expect from these
parties.  I think it is interesting to note that this was coming out of Margaret Thatcher's
tenure as Prime Minister which ended in 1990, and she was the first female PM and a
member of the Conservative party, so I was curious if her leadership would bring female
Conservative participation up to a level even with the other parties.  Also a random
note: the Green party in the UK has only ever had one member of parliament elected
to the House of Commons, who is female.  Her name is Caroline Lucas and she's been an mp
since 2010, so it is interesting that Green's have the highest expected proportion of female participation here.


Additional:

1)
We can conduct a goodness-of-fit test

2)
```{r}
c(Df = df.residual(binom.logit),
  Resid.dev = deviance(binom.logit),
  Pearson.X2 = sum(resid(binom.logit, type = "pearson")^2))
```
These are different, but I wouldn't feel comfortable saying they are significantly different.
We want these 3 numbers to be close, and I would say they are sufficiently close enough to
not worry about overdispersion.

3)
To handle any overdispersion, we could shift from using a binomial model to a 
quasi-binomial model.







POISSON:

Description:

Mean-Variance Relationship:

  The mean equals the variance in a poisson distribution

Dispersion:

  The dispersion parameter equals 1 for the poisson

Potential Link Functions:
Methods of Inference:
Saddlepoint Approximation and Central Limit Theorem:

  The saddlepoint approximation for the poisson will be sufficient when
    min(y) >= 3
  The central limit theorem will be sufficiently accurate for the poisson when
    min(y) >= 5

When to adjust for overdispersion:
What type of residual to use:
Diagnostics:
Scale transformations to use:
Primary effect size estimates:




Analysis:



Additional:




GAMMA:

Description:

Mean-Variance Relationship:

  The variance of mu for the gamma edm is mu squared

Dispersion:

  Unless y follows a normal or exponential distribution, the dispersion parameter
    will have to be estimated.  Thus, the likelihood ratio tests are based on
    F-tests

Potential Link Functions:
Methods of Inference:
Saddlepoint Approximation and Central Limit Theorem:

  The saddlepoint approximation for the gamma will be sufficiently accurate when
    phi <= 1/3
  The central limit theorem for the gamma will be sufficiently accurate when
    phi <= 1/5

When to adjust for overdispersion:
What type of residual to use:
Diagnostics:
Scale transformations to use:
Primary effect size estimates:




Analysis:



Additional:






