---
title: "Michael Hornsby EDRM 878 Assignment 2 Notebook"
output: html_notebook
---

```{r include=FALSE}
library(here)
library(tidyverse)
library(ggrepel)
library(scales)
library(readxl)
library(purrr)
library(dplyr)
library(car)
library(GLMsData)
library(MASS)
```


Problem Set 3.16


```{r}
data(blocks)

blocks$Trial <- as.factor(blocks$Trial)
blocks$Shape <- as.factor(blocks$Shape)

summary(blocks)
```


The model I used in 2.18:

```{r}
anova(lm(Time ~ Shape, data = blocks))
```



```{r}
blocks %>% 
  ggplot(aes(x = Shape, y = Time)) +
  geom_boxplot() +
  labs(title = "Boxplots of Time by Block Shape",
       x = "Block Shape",
       y = "Time")

my.model <- lm(Time ~ Shape, data = blocks)

resid.raw <- resid(my.model)

fit.values <- fitted(my.model)

plot(fit.values, resid.raw)
```

Looking at the spread of the boxplots and the outliers present, it isn't
exactly clear I can safely assume equal variance among the groups.  This is
supported by the plot of raw residuals.  The spread of residuals doesn't look
equal among the two groups.  One group in the residuals clearly has a larger
spread than the other, outliers included.

```{r}
influence.measures(my.model)
```

Looking at the influence measures, observations 6 and 76 are cause for concern.
Let's see if this is remedied in our later model.


Let's look at a qq-plot of time to see if a transformation is needed.

```{r}
blocks %>%
  ggplot() +
  geom_qq(aes(sample = Time)) +
  labs(title = "QQ-plot for Time")
```


This does show a pretty significant deviation from normality.
Let's look at a boxcox ladder of powers plot.

```{r}
boxcox(Time ~ Shape,
       lambda = seq(-2.00, 1.00, by = .05), data = blocks)
```

Based on this plot. I'm going to look at two transformations.
The first will be a log transformation of the Time response.
The second will be Time raised to the (-2/5).


```{r}
blocks$logTime <- log(blocks$Time)
blocks$transTime <- (blocks$Time)^(-2/5)
```

```{r}
blocks %>%
  ggplot() +
  geom_qq(aes(sample = logTime)) +
  labs(title = "QQ-plot for logTime")

blocks %>%
  ggplot() +
  geom_qq(aes(sample = transTime)) +
  labs(title = "QQ-plot for transTime")
```

Looking at these, I think I am going to stick with the (-2/5) transformation.
I know that transformation might be a little tricker to interpret or explain,
but I really think the qq-plot looks better.


```{r}
blocks %>% 
  ggplot(aes(x = Shape, y = transTime)) +
  geom_boxplot() +
  labs(title = "Boxplots of transTime by Block Shape",
       x = "Block Shape",
       y = "transTime")
```

The boxplots even look like they have equal variance this time!


Let's look at the new model.

```{r}
anova(lm(transTime ~ Shape, data = blocks))
```


```{r}
my.new.model <- lm(transTime ~ Shape, data = blocks)
```


```{r}
resid.raw <- resid(my.new.model)

fit.values <- fitted(my.new.model)

plot(fit.values, resid.raw)

influence.measures(my.new.model)
```

The boxplots look as though they have equal variance, the residual plot looks
much better and shows a fairly equal spread among the two groups, and the 
problem of outliers looks to be solved.  Surprisingly, it looks as though
the transformation of the Time response variable remedied the obvious
problems I had.  When we look at the influence measures, we have
even more observations marked as influential.  However, looking at the
residuals, I don't think these are a cause for concern here.  Given our sample
size and how reasonable everything seems looking at the residuals, I would be
extremely hesitant to simply remove any observations with high influence from
either model without further information.

***

**Very nice work!**

**20/20**
