---
title: "Michael Hornsby EDRM 878 Assignment 1 Notebook"
output: html_notebook
---

```{r include=FALSE}
library(here)
library(tidyverse)
library(ggrepel)
library(scales)
library(readxl)
library(purrr)
library(dplyr)
library(car)
library(GLMsData)
```


Problem Set 1.9

Part 1)

```{r}
data(blocks)

blocks$Trial <- as.factor(blocks$Trial)
blocks$Shape <- as.factor(blocks$Shape)

summary(blocks)
```



Part 2)

```{r}
blocks %>% 
  ggplot(aes(x = Shape, y = Time)) +
  geom_boxplot() +
  labs(title = "Boxplots of Time by Block Shape",
       x = "Block Shape")
```

```{r}
blocks %>% 
  ggplot(aes(x = Trial, y = Time)) +
  geom_boxplot() +
  labs(title = "Boxplots of Time by Trial")
```

```{r}
blocks %>% 
  ggplot(aes(x = Age, y = Time)) +
  geom_point() +
  labs(title = "Scatterplot of Time by Age")
```



Part 3

Looking at the boxplots of time taken by block shape, it appears as though
using cube blocks resulted in an overall longer time taken. 
I think I would have expected the opposite.  I would have guessed children
would be quicker at stacking cubes than cylinders.
Looking at the boxplots of time by trial, it looks as though children took
about the same time for both trials.  As a general trend in the scatterplot,
there doesn't seem to be an effect of age on time.  If I ignore those 2 points
greater than a time of 125, there doesn't seem to be an significant
relationship that I can visualize form the scatterplot.



Part 4

```{r}
blocks %>% 
  ggplot(aes(x = Shape, y = Number)) +
  geom_boxplot() +
  labs(title = "Boxplots of Number of Blocks by Block Shape",
       x = "Block Shape",
       y = "Number of Blocks")
```

```{r}
blocks %>% 
  ggplot(aes(x = Trial, y = Number)) +
  geom_boxplot() +
  labs(title = "Boxplots of Number of Blocks by Trial",
       y = "Number of Blocks")
```

```{r}
blocks %>% 
  ggplot(aes(x = Age, y = Number)) +
  geom_point() +
  labs(title = "Scatterplot of Number of Blocks by Age",
       y = "Number of Blocks")
```



Part 5)

Looking at the boxplots of Number by Shape, it appears as though children using
cubes used more blocks than children using cylinders on a fairly consistant
basis.  Children at trial 2 appear to use slightly more blocks than at trial 1,
although, I wouldn't think this effect would be significant at any reasonable
alpha level looking at the boxplots.
Looking at the scatterplot of number of blocks by age, there does appear to be
a slightly positive linear relationship between number of blocks and age.  As
age increases, the number of blocks used on average appears to increase
slightly.



Problem Set 2.18)

Part 1)
