---
title: "Michael Hornsby EDRM 878 Assignment 4"
output: html_notebook
---

```{r include=FALSE}
library(here)
library(tidyverse)
library(ggrepel)
library(scales)
library(readxl)
library(purrr)
library(dplyr)
library(car)
library(GLMsData)
library(MASS)
library(statmod)
```


Import and tidy data:

```{r}
data(blocks)

blocks$Trial <- as.factor(blocks$Trial)
blocks$Shape <- as.factor(blocks$Shape)

summary(blocks)
```


My 6.10 glm:

```{r}
summary(glm(Number ~ Age, family = poisson(link = log), data = blocks))
```

Problem 7.4

The Wald test for age (same as above):


```{r}
summary(glm(Number ~ Age, family = poisson(link = log), data = blocks))
```

This does suggest age makes a significant predictor of the number of blocks when using the Poisson regression for the Number of Blocks response variable.



The score test:

```{r}
blocks.m0 <- glm(Number ~ 1, data = blocks, family = poisson(link = "log"))
z.stat <- glm.scoretest(blocks.m0, blocks$Age)
p.val <- 2 * pnorm(abs(z.stat), lower.tail = FALSE)
round(c(score.stat = z.stat, P = p.val), 4)
```

The score test also suggests age makes for a significant predictor with just about the same p-value.



The likelihood ratio test:

```{r}
fit <- glm(Number ~ Age, data = blocks, family = poisson(link = "log"))
as.data.frame(anova(fit, test = "Chisq"))
```

The likelihood ratio test p-value is similar as well.



The results from all 3 tests suggest the same conclusion.  All 3 tests results in p-values very close to 0.008, which is significant at any reasonable alpha level.  Thus, all 3 tests conclude that age is a significant predictor for the number of blocks when using a poisson distribution and a log link for the distribution of the number of blocks response.


```{r}
min(blocks$Number)
```


The saddlepoint approximation is sufficiently accureate when the min(yi) is greater than or equal to 3 when using the poisson distribution.  Thus, the saddlepoint approximation should be sufficiently accurate as our min(yi) = 3.



The Central Limit Theorem is sufficiently accurate when the min(yi) is greater than or equal to 5 when using the poisson distribution.  Thus, we cannot expect the Central Limit Theorem to be sufficiently accurate as our min(yi) = 3.


95% Wald confidence intervals for regression coefficients:

```{r}
wald.fit <- glm(Number ~ Age, family = poisson(link = log), data = blocks)
printCoefmat(coef(summary(wald.fit)))
confint(wald.fit)
```


Plot of relationship:

```{r}

plot(Number ~ Age, data = blocks)

abline(a = exp(1.34470), b = exp(0.14151))
abline(a = exp(0.9024058), b = exp(0.0377308), lty = 2)
abline(a = exp(1.7742235), b = exp(0.2471074), lty = 2)
```



Problem 8.11:

Once again, the model:

```{r}
my.model <- glm(Number ~ Age, family = poisson(link = log), data = blocks)
```

Let's see if I've chosen the right EDM, my residuals should have a normal distribution:

```{r}
qr.my.model <- qresid(my.model)
qqnorm(qr.my.model, las = 1)
qqline(qr.my.model)
```

This looks ok, not spectacular.  I would think it shows that the tails are a little light, but I don't think this would invalidate the model.


Check for outliers:

```{r}
rs <- cbind(rD = resid(my.model),
            "r'D" = rstandard(my.model),
            "r''" = rstudent(my.model),
            rQ = qresid(my.model))

head(rs)

apply(abs(rs), 2, max)
```

These maximum residuals look pretty good I would say, nothing extreme.

Let's check for influential observations:

```{r}
im <- influence.measures(my.model)
names(im)
im$infmat <- round(im$infmat, 3)
head(im$infmat)
colSums(im$is.inf)
```

```{r}
my.model.cd <- cooks.distance(my.model)

plot(my.model.cd,
     type = "h",
     ylab = "Cook's distance",
     las = 1)

plot(dffits(my.model),
     type = "h",
     las = 1,
     ylab = "DFFITS")

infl <- which.max(my.model.cd)
infl
my.model.cd[infl]
```

All of these values look well within reasonable bounds.  I don't think I'm concerned about any of these Cook's distance values or DFFITS.

In summary, my residuals look good, I don't see any outliers, and I don't see any observations with unacceptably high influence or weight.